Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}

### 2021-11-22 article.tex
update paper

The predictive distribution of a probabilistic forecast can be represented in different ways with implications for the appropriate evaluation approach. For most forecasting problems, predictive distributions are not readily available in a closed form (and the \pkg{scoringutils} package therefore does not support scoring them directly). Instead, predictive distributions are often represented by a set of quantiles or predictive samples. Predictive samples require a lot of storage space and also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution. For that reason, often quantiles or central prediction intervals are reported instead [citation FORECAST HUBS]. For binary prediction targets, common in many classification problems, a probabilistic forecasts is represented by the probability that an outcome will come true. DO I NEED TO TALK ABOUT MULTINOMIAL CLASSIFICATION PROBLEMS? Table \ref{tab:forecast-types} summarises the different forecast types and formats. The general forecasting paradigm \cite{gneitingProbabilisticForecastsCalibration2007}, however, that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.

\begin{table}[h!]
\centering
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{4.5cm}}
\toprule
\textbf{Forecast type} & \textbf{Target type} & \textbf{Representation of the predictive distribution}\\
\midrule
Point forecast & continuous \newline  integer \newline  binary & one single number for the predicted outcome\\
\cmidrule{1-3}\pagebreak[0]
 & continuous \newline  integer & predictive samples \newline   quantiles \newline   closed analytical form\\
\cmidrule{2-3}\nopagebreak
\multirow{-2}{3.5cm}{\raggedright\arraybackslash Probabilistic forecast} & binary & binary probabilities\\
\bottomrule
\end{longtable}\caption{\label{tab:forecast-types} Ssummary of the different forceast types and forecast formats}
\end{table}

@@

Some of the metrics in \pkg{scoringuitls} focus only on sharpness or on calibration. Others, called proper scoring rules, combine both aspects into a single number. The former can be helpful to learn about specific model aspects and improve them, the latter are especially useful to assess and rank predictive performance of a forecaster. The following gives an introduction to how these metrics can be used to evaluate forecasts. Table \ref{tab:table-summary-scores} shows an overview of the metrics implemented in \pkg{scoringutils}. Table \ref{tab:score-table} in the Appendix gives a thorough explanation of all the metrics.

\begin{table}[h!]
\centering
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{4.5cm}>{\raggedright\arraybackslash}p{4cm}}
\toprule
Metric & Target types & Forecast formats & Properties\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
Metric & Target types & Forecast formats & Properties\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{CRPS (Continuous) ranked probability score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
\addlinespace
Log score & continuous, (integer not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
\addlinespace
\cellcolor{gray!6}{WIS (Weighted) interval score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
\addlinespace
DSS Dawid-Sebastiani score & continuous, integer & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
\addlinespace
\cellcolor{gray!6}{Brier score} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
\addlinespace
Interval coverage & continuous, integer & interval forecasts (needs matching quantiles) & measure for calibration\\
\addlinespace
\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
\addlinespace
Probability integral transform (PIT) & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
\addlinespace
\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
\addlinespace
Bias & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
\addlinespace
\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{compares performance of two models}\\
\addlinespace
Relative skill & depends on scored & depends on score & Ranks models based on pairwise comparisons\\*
\end{longtable}\caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
\end{table}

\subsection{Proper scoring rules}

Proper scoring rules \citep{gneitingStrictlyProperScoring2007} jointly assess sharpness and calibration and assign a single numeric value to a forecast. A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average. This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state their best estimate.

For sample-based forecasts, the (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], and the Dawid-Sebastiani-score (dss) [CITATION] are available in \pkg{scoringutils}. These are implemented as wrappers around functions from the \pkg{scoringRules} package (which also has closed-form versions of the scoring rules available). They are in principle applicable to continuous as well as integer forecasts. The \pkg{scoringRules} implementation of the log score, however, requires a kernel density estimation that may be inappropriate for integer values (see also Table \ref{tab:table-summary-scores}) and is therefore not available for discrete predictions in \pkg{scoringutils}. 
For forecasts in an interval or quantile format, the weighted interval score (wis) \citep{bracherEvaluatingEpidemicForecasts2020a} is available. For an increasing number of equally-spaced prediction intervals, the wis converges to the crps and therefore has very similar properties. Binary forecasts can be scored using the Brier score (bs) [CITATION].

%sample-based proper scoring rules\usepackage{multirow}
When scoring forecasts in a sample-based format, the choice is usually between the log score and the crps. The dss is much less commonly used. It is easier to compute, but apart from that does not have immediate advantages over the former two. Crps and log score differ in two important aspects: the first is sensitivity to distance \cite{winklerScoringRulesEvaluation1996}, the second is how harshly far-off predictions are punished.

The crps is a so-called global scoring rule, which means that the entire predictive distribution is taken into account when scoring a single forecast. The log score, on the other hand is local. The resulting score does not depend on the overall distance between the observed value and the distribution, but only on the probality density assigend to the actual outcome. Imagine two forecasters, A and B, who forecast the number of points scored in a basketball game. If both forecasters assigned the same probability to the true outcome (100 points), but A assigned higher probability to extreme outcomes far away from the actually observed outcome, then A will receive a worse score than B. The log score, in contrast, is a local scoring rule that only scores the probability assigned to the actual outcome and ignores the rest of the predictive distribution. Judged by the log score, A and B would receive exactly the same score. Sensitivity to distance (taking the entire predictive distribution into account) is arguably an advantage in most settings that involve decision making. Forecaster A's prediction that assigns high probability to results far away from the observed value is arguably less useful than B's forecast that assigns higher probability to values closer to it (the probability assigned to the actual outcome being equal for both forecasts). The log score is only implicitly sensitive to distance if we assume that values close to the observed value are actually more likely to occur. It may, however, be more appropriate for inferential purposes (see \cite{winklerScoringRulesEvaluation1996}).

% stable vs. non-stable
A second important difference is how forecasts are treated that deviate strongly from the observered outcome. The crps can be thought of as a generalisation of the absolute error to a predictive distribution. It therefore scales linearly with the distance between forecast distribution and true value. The log score, however, is the log of the predictive density evaluated at the observed value. It can therfore quickly go to negative infinity if the probability assigned to the observed outcome is close to zero. The crps is therefore considered more stable than the log score. The behaviour of the dss is in between the two. Whether or not harsh punishment of bad predictions is desirable or not depends of course on the setting. \cite{bracherEvaluatingEpidemicForecasts2020a} exemplify that in practice there may indeed be substantial differences between how the crps and log score judge the same forecast.

% properties of the WIS + decomposition
For quantile forecasts the wis is the best available option and can be understood as an approximation to the crps. One additional benefit of the wis is that it can easily be decomposed into three additive components: an uncertainty penalty (sharpness) for the width of a prediction interval and penalties for over- and underprediction (if a value falls outside of a prediction interval). This can be very helpful in diagnosing model problems. It may therefore even be useful to convert samples into quantiles and use the wis instead of the crps to make use of this decomposition for the purpose of model diagnostics. 
% fact that scores scale with the overall level of the prediction
Both crps and wis, as generalisations of the absolute value, suffer from the fact that overall scores depend on the order of magnitude of the quantity we try to forecast. It can therefore be difficult to compare forecasts for very different targets. One possibility to address this is the use of pairwise comparisons introduced later. In general, this is much less of a problem for the log score and the dss. IS IT ONE AT ALL?

@@

In addition to the proper scoring rules outlined above, \pkg{scoringutils} makes numerous metrics available to evaluate calibration and sharpness independently. This is especially helpful for model diagnostics.

@@

Several strategies have been proposed to detect systematic deviations of the predictive distributions from the observations (see e.g. \cite{funkAssessingPerformanceRealtime2019}; \cite{gneitingProbabilisticForecastsCalibration2007}; \cite{gneitingStrictlyProperScoring2007}). Using \pkg{scoringutils}, we can look at three different aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).

@@

% interval coverage and quantile coverage
Another way to look at calibration (precisely: probabilistic calibration in \cite{gneitingProbabilisticForecastsCalibration2007}) is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution. This is most easily understood in the context of quantile forecasts, but can easily be transferred to sample-based continuous and integer forecasts as well. 
%
To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution. If the 0.05, 0.25, 0.75, and 0.95 quantiles are given, then 50\% of the true values should fall between the 0.25 and 0.75 quantiles and 90\% should fall between the 0.05 and 0.95 quantiles. We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated. 
%
To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution. This allows to diagnose issues in the lower and upper tails of the prediction intervals separately. A similar way to visualise the same information is a PIT histogram. In order to conveniently assess deviations between the predictive distribution and the true data-generating distribution we can transform the observed values using the probability integral transformation (PIT) \citep{dawidPresentPositionPotential1984} (see more details in Table \ref{tab:score-table}). If both distributions are equal, the transformed values will follow a uniform distribution. A histogram of the transformed values can help to diagnose systematic differences between the predictions and the observed values. Figure \ref{fig:calibration-plots} exemplifies the characteristic shape of certain systematic deviations of the predicitve distribution from the true data-generating distribution. In the PIT histograms, bias leads to a triangular shape, overdispersion results in a hump shaped form and underdispersion in a U-shape. ADD INTERPRETATION FOR QUANTILE AND INTERVAL COVERAGE PLOTS HERE.

\begin{figure}[h]
\centering
\includegraphics{plots/calibration-diagnostic-examples.png}
\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
\end{figure}

@@

% ==============================================================================
\section{Evaluating UK short-term forecasts}
\subsection{The data}

To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020. Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M). The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic. Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE). The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL].

- timing (weekly?) and number of forecast dates
- the four targets
- the models. 
- is the set complete?

We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
% 
\begin{Schunk}
\begin{Sinput}
R> # install and load data pacakge from external repository
R> # remotes::install_github("sbfnk/covid19.forecasts.uk")
R> 
R> # load packages
R> library(covid19.forecasts.uk)
R> library(dplyr)
R> library(scoringutils)
R> # load truth data
R> data(covid_uk_data)
R> # head(covid_uk_data)
R> 
R> # load forecasts
R> data(uk_forecasts)
R> # head(uk_forecasts)
\end{Sinput}
\end{Schunk}
% 
Let us take a first look a the data:
% 
\begin{Schunk}
\begin{Sinput}
R> glimpse(covid_uk_data)
\end{Sinput}
\begin{Soutput}
Rows: 4,174
Columns: 6
$ geography  <fct> London, London, London, London, London, London, L…
$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospita…
$ value_desc <fct> Hospital admissions, Hospital admissions, Hospita…
$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0…
$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23, …
$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546, …
\end{Soutput}
\begin{Sinput}
R> glimpse(uk_forecasts)
\end{Sinput}
\begin{Soutput}
Rows: 1,254,513
Columns: 8
$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, E…
$ geography     <chr> "East of England", "East of England", "East of…
$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hosp…
$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-3…
$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-0…
$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05…
$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, 0…
$ value_desc    <fct> Hospital admissions, Hospital admissions, Hosp…
\end{Soutput}
\end{Schunk}
% 
To bring the forecasts into the format needed for the evaluation, some minor changes need to be made to the data. The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}. While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public. The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way. 
% 
\begin{Schunk}
\begin{Sinput}
R> uk_forecasts <- rename(uk_forecasts, prediction = value)
R> covid_uk_data <-rename(covid_uk_data, true_value = value)
R> combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
\end{Sinput}
\end{Schunk}
% 
Before we start with scoring the forecasts, it makes sense to start the evaluation process by visualising the data. To get a feeling for how complete the data set is, we can run the following code to obtain a heatmap with the number of available forecasts: 
% 
\begin{figure}[h]
\centering
\begin{Schunk}
\begin{Sinput}
R> show_avail_forecasts(combined, 
+                       x = "creation_date", 
+                       show_numbers = FALSE,
+                       legend_position = "bottom",
+                       facet_formula = ~ value_desc)
\end{Sinput}
\end{Schunk}
\includegraphics{plots/plot-show-availability}
\caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
\end{figure}
%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
% 
Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance. By default, the function treats a set of different quantiles or samples as one forecast. However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts.

Forecasts can be visualised using the \fct{plot\_predictions} function. The function accepts either a single combined data set or two separate truth data sets that can be merged together. Forecasts and observed values can be filtered independently, to show for example only predictions from a certain model or a specific number of weeks of true data before the forecast. Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data. To obtain, for example, a specific forecast from a model we are interest in, we can call: 
% 
\begin{figure}[h!]
\centering
\begin{Schunk}
\begin{Sinput}
R> locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
R> plot_predictions(truth = covid_uk_data,
+                   forecasts = uk_forecasts,
+                   filter_both = list(paste("geography %in%", locations, collapse = "")),
+                   filter_truth = list('value_date <= "2020-06-22"',
+                                       'value_date > "2020-06-01"'), 
+                   filter_forecasts = list('model == "SIRCOVID"', 
+                                           'creation_date == "2020-06-22"'),
+                   x = "value_date",
+                   facet_formula = geography ~ value_type) + 
+    ggplot2::theme(legend.position = "bottom")
\end{Sinput}
\end{Schunk}
\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
\end{figure}
% 
The output is shown in Figure \ref{fig:forecast-visualisation}.

\subsection[Scoring forecasts with eval_forecasts]{Scoring forecasts with \fct{eval\_forecasts}}

>>>>

The predictive distribution of a probabilistic forecast can be represented in different ways with implications for the appropriate evaluation approach. For most forecasting problems, predictive distributions are not readily available in a closed form (and the \pkg{scoringutils} package therefore does not support scoring them directly). Instead, predictive distributions are often represented by a set of quantiles or predictive samples. Predictive samples require a lot of storage space and also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution. For that reason, often quantiles or central prediction intervals are reported instead [citation FORECAST HUBS]. For binary or multinomial prediction targets, common in many classification problems, a probabilistic forecasts is represented by the probability that an outcome will come true. Table \ref{tab:forecast-types} summarises the different forecast types and formats. While specific metrics may differ depending on the forecast type or format, the general forecasting paradigm \cite{gneitingProbabilisticForecastsCalibration2007} that guides the evaluation process is the same.

\begin{table}[h!]
\centering
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{4.5cm}}
\toprule
\textbf{Forecast type} & \textbf{Target type} & \textbf{Representation of the predictive distribution}\\
\midrule
Point forecast & continuous \newline  discrete \newline  binary & one single number for the predicted outcome\\
\cmidrule{1-3}\pagebreak[0]
 & continuous \newline  discrete & predictive samples \newline   quantiles \newline   closed analytical form\\
\cmidrule{2-3}\nopagebreak
\multirow{-2}{3.5cm}{\raggedright\arraybackslash Probabilistic forecast} & binary & binary probabilities\\
\bottomrule
\end{longtable}\caption{\label{tab:forecast-types} Forecasts can be distinguished by whether they are probabilistic in nature, or a point forecast only. Depending on the type of the target (discrete, continuous or binary) different representation of the predictive distribution are possible.} 
\end{table}

@@

Some of the metrics in \pkg{scoringuitls} focus only on sharpness or on calibration. Others, called proper scoring rules, combine both aspects into a single number. The former can be helpful to learn about specific model aspects and improve them, the latter are especially useful to assess and rank predictive performance of a forecaster. The following gives an introduction to how these metrics can be used to evaluate forecasts. Table \ref{tab:table-summary-scores} shows an overview of the metrics implemented in \pkg{scoringutils}. Table \ref{tab:score-table} in the Appendix gives a mathematical definition and thorough explanation of all the metrics. Figure \ref{fig:calibration-plots} gives an applied example of different scoring metrics and visualisations for simple toy data.

% \begin{table}[h!]
% \centering

% \caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
% \end{table}

\newpage
\begin{table}[h!]
\centering
\begingroup\fontsize{10}{12}\selectfont

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.0 cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.2cm}>{\raggedright\arraybackslash}p{4.5cm}}
\toprule
Metric & Discrete & Continuous & Binary & Closed-form & Samples (approx.) & Quantiles & Properties\\
\midrule
\cellcolor{gray!6}{(Continuous) ranked probability score (CRPS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
\addlinespace
Log score (logS) & ($\checkmark$) & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, local, unstable for outliers\\
\addlinespace
\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK AGAIN} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
\addlinespace
Dawid-Sebastiani score (DSS) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
\addlinespace
\cellcolor{gray!6}{Brier score (BS)} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule}\\
\addlinespace
Interval coverage & $\checkmark$ & $\checkmark$ &  &  & THINK AGAIN & $\checkmark$ & measure for calibration\\
\addlinespace
\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK AGAIN} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measure for calibration}\\
\addlinespace
Probability integral transform (PIT) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & assesses calibration\\
\addlinespace
\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
\addlinespace
Bias & $\checkmark$ & $\checkmark$ & $\checkmark$??? & $\checkmark$ & $\checkmark$ & $\checkmark$ & captures tendency to over-or underpredict (aspect of calibration)\\
\addlinespace
\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{compares performance of two models}\\
\addlinespace
Relative skill & depends & depends & depends & depends & depends & depends & Ranks models based on pairwise comparisons\\
\bottomrule
\end{longtable}
\endgroup{}\caption{\label{tab:table-summary-scores2} Summary table of scores available in scoringutils}
\end{table}

\begin{figure}[h]
\centering

\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
\end{figure}

@@

Evaluating calibration and sharpness independently is helpful for model diagnostics. To that end \pkg{scoringutils} makes numerous metrics available that aim to capture different aspects of sharpness and calibration.

@@

Calibration means consistency between forecasts and observed values, but there are various ways in which a forecast can systematically deviate from the observations (see \cite{gneitingProbabilisticForecastsCalibration2007} for a discussion of different forms of calibration MAYBE IT WOULD BE A GOOD IDEA TO INCLUDE THESE DIFFERENT FORMS OF CALIBRATION HERE).
\pkg{scoringutils} allows the user to examine three different sub-aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).

@@

% interval coverage and quantile coverage
Another way to look at calibration (precisely: probabilistic calibration in \cite{gneitingProbabilisticForecastsCalibration2007}) is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution. This is most easily understood in the context of quantile forecasts, but can easily be transferred to sample-based continuous and discrete forecasts as well. 
%
To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution. If the 0.05, 0.25, 0.75, and 0.95 quantiles are given, then 50\% of the true values should fall between the 0.25 and 0.75 quantiles and 90\% should fall between the 0.05 and 0.95 quantiles. We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated. 
%
To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution. This allows to diagnose issues in the lower and upper tails of the prediction intervals separately. A similar way to visualise the same information is a PIT histogram. In order to conveniently assess deviations between the predictive distribution and the true data-generating distribution we can transform the observed values using the probability integral transformation (PIT) \citep{dawidPresentPositionPotential1984} (see more details in Table \ref{tab:score-table}). If both distributions are equal, the transformed values will follow a uniform distribution. A histogram of the transformed values can help to diagnose systematic differences between the predictions and the observed values. Figure \ref{fig:calibration-plots} exemplifies the characteristic shape of certain systematic deviations of the predicitve distribution from the true data-generating distribution. In the PIT histograms, bias leads to a triangular shape, overdispersion results in a hump shaped form and underdispersion in a U-shape. ADD INTERPRETATION FOR QUANTILE AND INTERVAL COVERAGE PLOTS HERE.

@@

% ==============================================================================
\section{Evaluating UK short-term forecasts}

The following section shows an example evaluation of short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020. Forecasts were produced by six research groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M). The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic. Predictions submitted to SPI-M were aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE). All predictions as well as most of the observed values are now publicly available and discussed in more depth in [FUNK ET AL]. 
% - timing (weekly?) and number of forecast dates
% - the four targets
% - the models.

The evaluation process looks as follows: First we load, prepare and visualise (some of) the data. Then we obtain forecast scores and a model ranking based on pairwise-comparisons, followed by a more detailed analysis of calibration and sharpness.

\subsection{Preparing and visualising the data}

We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
% 
\begin{Schunk}
\begin{Sinput}
R> # install and load data pacakge from external repository
R> # remotes::install_github("sbfnk/covid19.forecasts.uk")
R> 
R> # load packages
R> library(covid19.forecasts.uk)
R> library(dplyr)
R> library(scoringutils)
R> # load truth data
R> data(covid_uk_data)
R> # head(covid_uk_data)
R> 
R> # load forecasts
R> data(uk_forecasts)
R> # head(uk_forecasts)
\end{Sinput}
\end{Schunk}
% 
Let us take a first look a the data:
% 
\begin{Schunk}
\begin{Sinput}
R> glimpse(covid_uk_data)
\end{Sinput}
\begin{Soutput}
Rows: 4,174
Columns: 6
$ geography  <fct> London, London, London, London, London, London, L…
$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospita…
$ value_desc <fct> Hospital admissions, Hospital admissions, Hospita…
$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0…
$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23, …
$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546, …
\end{Soutput}
\begin{Sinput}
R> glimpse(uk_forecasts)
\end{Sinput}
\begin{Soutput}
Rows: 1,254,513
Columns: 8
$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, E…
$ geography     <chr> "East of England", "East of England", "East of…
$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hosp…
$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-3…
$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-0…
$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05…
$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, 0…
$ value_desc    <fct> Hospital admissions, Hospital admissions, Hosp…
\end{Soutput}
\end{Schunk}
% 
Some minor changes need to be made to the data before we can evaluate the forecasts. The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}. While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public. The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way. 
% 
\begin{Schunk}
\begin{Sinput}
R> uk_forecasts <- rename(uk_forecasts, prediction = value)
R> covid_uk_data <-rename(covid_uk_data, true_value = value)
R> combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
\end{Sinput}
\end{Schunk}
% 
It is helpful to start the evaluation process by visualising the available data, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance. The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts. By default, the function treats a set of different quantiles or samples as one forecast. However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts. 
% 
\begin{figure}[h]
\centering
\begin{Schunk}
\begin{Sinput}
R> show_avail_forecasts(data = combined, 
+                       x = "creation_date", 
+                       show_numbers = FALSE,
+                       legend_position = "bottom",
+                       facet_formula = ~ value_desc)
\end{Sinput}
\end{Schunk}
\includegraphics{plots/plot-show-availability}
\caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
\end{figure}
%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
% 
The forecasts and observed values themselves can be visualised using the \fct{plot\_predictions} function. Data visualisation is of course highly context-dependent, but the function tries to accommodate most practical use-cases. In addition to basic plotting functionality it offers the user an optional ad hoc way to filter both forecasts and observed values. This makes it possible to tweak the plot in a beginner-friendly way without having to manipulate the data separately. Forecasts and observed values can be passed in separately (and are merged internally) or as a single data.frame. Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data. To display, for example, death forecasts from the SIRCOVID model made on the 22nd July 2020 for the four UK countries and add three weeks of prior observations to the plot, we can call: % 
\begin{figure}[h!]
\centering
\begin{Schunk}
\begin{Sinput}
R> locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
R> plot_predictions(truth = covid_uk_data,
+                   forecasts = uk_forecasts,
+                   filter_both = list(paste("geography %in%", locations, collapse = ""), 
+                                      'value_type == "death_inc_line"'),
+                   filter_truth = list('value_date <= "2020-06-22"',
+                                       'value_date > "2020-06-01"'), 
+                   filter_forecasts = list('model == "SIRCOVID"', 
+                                           'creation_date == "2020-06-22"'),
+                   x = "value_date",
+                   facet_formula = ~ geography) + 
+    ggplot2::theme(legend.position = "bottom")
\end{Sinput}
\end{Schunk}
\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
\end{figure}
% 
The output is shown in Figure \ref{fig:forecast-visualisation}.

\subsection[Scoring forecasts and obtaining a model ranking]{Scoring forecasts with \fct{eval\_forecasts}}

The actual scoring of forecasts based on observed values can be performed using the function \fct{eval\_forecasts}. The function automatically detects the forecast type and format, applies the appropriate scoring metrics and aggregates results as desired by the user. Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation. The function requires a \code{data.frame} or similar with some pre-specified columns (e.g. a column called "prediction", one called "true\_value" and more depending on the forecast format). Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons. To illustrate the requirements, \pkg{scoringutils} provides example data for all possible forecast formats. It also offers functions to to transform between various formats, e.g. from a sample based format to a quantile format.
